\Large \bf What Does the Weather Say about Yields in Kenya

### 2018-11-14 writing/draft1/manuscript2.tex
adding

According to scientific consensus, climate change does exist and it is human caused \citep{oreskes2004, mccarthy2001impacts, Cook2013COnsensus}. However, public opinions about climate change are far away from consensus \citep{leiserowitz2012, Pew2012}. Without taking actions to prevent or mitigate climate change and its consequences, the effects of global warming could be disastrous \citep{IPCC5WG1, IPCC5, Seneviratne2012}.

Public attitudes towards natural hazards and risk perception are important drivers of policy decision making (\citealp{BritishColumbia, bookDisastersPrep}). Whether and how well climate change will be tackled depends on public opinion heavily. But what are the main factors influencing climate change perception and awareness among general population? Large literature examines role of personal characteristics, demographics and behavioural variables in climate change awareness and risk perception by means of survey or experimental methods. Recent survey-based studies include \citet{Leiserowitz2015}, who exploits the Gallup World Poll data and conclude that civic engagement, communication access and education are the most important predictors of climate change awareness while beliefs about causes and perception of local temperature changes are main predictors of climate risk perception. Another example of a survey-based analysis of environmental attitudes is \citet{Morrison2015}, who use ordered logit models to investigate relationship between religion and climate change attitudes and behaviour. They conclude that Buddhists, atheists and agnostics are the most engaged with climate change while Christian literalists are the least engaged. Also \citet{Carlsson2013} use survey data to estimate preferences of redistribution of burden caused by CO\textsubscript{$2$} emissions in China and in the United States. Understandingly, both Chines and Americans prefer rules of redistribution which are less costly for their country. However, these rules differ for the two countries. US respondents prefer the current emissions rule while in China, the historical emissions rule is preferred.

As for papers based on experiments, for instance \citet{Braaten2014} analyses motivation of good feeling from giving in contrast to 'pure altruism' in climate change context. He concludes that 'warm glow' is important for motivating of environmentally friendly behaviour. Another example of a study which uses experimantal methods is \citet{GlenkColombo2013}, who conduct a choice experiment in Scotland. Based on the results, they claim that an outcome related risk is an important attribute in choice of land-based climate change mitigation project.

The above mentioned studies are just a small portion of recent papers focused on climate concerns and climate knowledge. The previous studies of public attitudes and knowledge about environment are usually based either on survey or on experimental methods. Surveys can be conducted over large, representative samples and experiments are powerful tools to infer parameters of utility functions, measures of risk and time preferences or social value orientation (\citealp{ifcher2011, murphy2011SVO, tanaka2010riskTime}). Each of these methods, however, suffers from serious drawbacks. Surveys often lead to hypothetical bias while experimental data tend to be affected by artificial settings and small, non-representative samples which often consist of students. We contribute by overcoming some of these shortcomings as we use a dataset which was created by surveying a large representative sample of respondents in an experimental, interactive and dynamic way. Experiments are usually computer based and their participants respond to various situations on a screen. We replicate this set-up in a live sample survey by including the experimental methods as a part of the survey \citep{SurveyUK}. The experimental set-up covers attitudes towards risk, attitudes towards equity including altruism and time preferences. We will explore effects of these attitudes on stances towards climate change and climate knowledge.

We also investigate influence of other characteristics on environmental attitudes including standard demographic data such as age, sex, race, ethnicity, religion, education, sector, occupation, date of birth, siblings, questions about assets, debts and family income. We further examine role of financial literacy and numeracy and we also investigate preferences regarding government spending and income redistribution, cultural and political ideology and world-view. As there is no general consensus on what are the main determinants of climate change perception and climate literacy, we start the explanatory analysis using a least absolute shrinkage and selection operator (lasso) to select significant predictors from almost $70$ candidates.

We further contribute by showing substantially different results from those of \citet{Newell2015}, who experimentally measure individual discount rates and analyse their role in energy efficiency decisions of US households. They find a negative and significant effect. More specifically, \citet{Newell2015} conclude that willingness to pay (WTP) for annual operating energy cost savings decrease in discount rates. This disagrees with our results as we do not find any evidence of significant effects of individual discount rates on any of our dependent variables, including WTP. We argue that our estimates are more precise than those of~\citet{Newell2015} as our sample size is substantially larger.\footnote{Our estimation samples include between $5659$ and $5749$ respondents while the estimation sample of~\citet{Newell2015} has $879$ observations.} We also cover much broader spectrum of potential predictors and we use more precise estimator, in concrete multisplit lasso with resampling.

Our additional contribution is a partial replication of~\citet{Kahan2015}. Consistently with his results, we find that climate knowledge measured by the 'ordinary climate science intelligence' (OCSI) instrument\footnote{For a more detailed description of the OCSI instrument see Section~\ref{ClimateVars}.} does not depend on measures of personal ideology and cultural world-view as opposed to other, previously used measures of climate knowledge (\citealp{Hamilton2011, Kahan2012, Kellstedt2008}). In accordance with \cite{Kahan2015}, our results show that climate knowledge is positively correlated with numeracy. We also detect association between our measure of climate knowledge and gender.

In accordance with previous research (\citealp{KahanEtAl2012, Kahan2015, Kellstedt2008}) we find that stated climate change risk perception does not increase with numeracy and financial literacy (we use these variables as proxies for ability of analytical reasoning and capacity to make use of quantitative information although they can also be interpreted as tests of the respondents' attentiveness during the survey) as one may intuitively assume \citep{Weber2011}. As a matter of fact, we find that individuals' concerns about climate change  decline as numeracy and financial literacy increase and it is also closely related to respondents' gender and cultural and ideological world-view. This is consistent with previous literature (\citealp{Kahan2015, Kellstedt2008, WHITMARSH2011}). We particularly show that the respondents, who agree with the statement that 'Government should redistribute income from the better off to those who are less well off' (we will further refer to this statement as 'government should redistribute income', 'degree of agreement with income redistribution' or simply 'income redistribution'), which we use as a measure of cultural or ideological world-view, are more likely to take climate change more seriously and have higher WTP for its mitigation than those not agreeing with this statement. Consistently with recent literature (e.g.~\citealp{Kahan2015,KahanEtAl2012,Hamilton2011, HamiltonKeim2009}), we find evidence suggesting that the ideological polarization over climate change is stronger among people who are more proficient in numeracy and comprehension of quantitative information.

We detect other significant predictors of WTP for climate change mitigation by means of gas and electricity tax. These are age, inequity aversion, perception of equality of intergenerational allocation of resources and risk assessment consistency. Expectedly, the respondents who consider themselves to be more affected by climate change than by climate policy have higher WTP than those who feel to be more affected by climate policy. Consistently with previous literature, we find negative and significant effect of age \citep{Hamilton2011, Kellstedt2008, Hayes2001gender}. The impacts of inequity aversion are mixed and WTP is significantly higher for those respondents who think that their income and standards of living are about the same compared to their parents (at the same stage of their lives) and their children (at the same stage of their lives as respondent currently is). We estimate an analogous model for WTP by means of transport fuel duty as a robustness test and the results are very comparable. The estimates are robust.

Perhaps surprisingly, we did not find the behavioural characteristics to be significant predictors of our climate perception or climate knowledge measures. The only exception is inequity aversion, which has significant effects on WTP for climate change mitigation.

The paper proceeds as follows. In Section~\ref{Method} we discuss methods, in particular multisplit lasso and jackknife ordinary least squares (OLS). Section~\ref{Data} describes the dataset used for our analysis and how the important variables were obtained. In Section~\ref{Results} we present and discuss the results. More specifically, Section~\ref{ResKnow} is focused on climate knowledge, in Section~\ref{ResPerc} we present estimates of climate seriousness perception models and Section~\ref{ResWTP} describes models for WTP. In Section~\ref{Robust} we estimate alternative specifications for each dependent variable to verify robustness of our results. We summarise our findings in Section~\ref{Summary} and in Section~\ref{Caveats} we discuss caveats. Section~\ref{Implications} includes policy implications and concluding remarks.

\FloatBarrier
\section{Methodology}\label{Method}

@@

For simplicity, we set $\gamma$ in~\eqref{pvals} to be equal to $0.5$ for every application of a multisplit lasso in this study. Each time we perform $H=100$ splits (we believe that this number is sufficient as \citet{pValsLasso} use $50$ sample splits per simulation) and we always use one third of the sample for the variable selection using lasso and the rest for the OLS estimation and obtaining \textit{p}-values.

A large fraction of our potential predictors are categorical variables because large part of the survey data was collected by multiple choice questions. However, the multisplit lasso selects individual predictors rather than groups of variables. Therefore, it can happen that a model specified by a multisplit lasso includes a dummy variable for one category of a particular categorical variable but it does not include dummy variables for its remaining categories. An obvious way how to overcome this issue would be to add the remaining dummy variables and use an \textit{F}-test to determine the joint significance of the group. If the \textit{F}-test implies that the categories are jointly significant, they should all stay in the model and they should be left out otherwise. However, the solution is not so straightforward with a multisplit lasso as it is not obvious on which subsample we should perform the \textit{F}-test. Therefore, for each model specified by a multisplit lasso, we decided to perform a following procedure which is sometimes called jackknife resampling. We will further refer to the procedure as a jackknife OLS. We again randomly split the dataset into two subsamples. The bigger subsample has size of two thirds of the original sample and it is used for OLS estimation and calculation of \textit{p}-values of the model with predictors selected by multisplit lasso.\footnote{Sample splitting can generally result in loss of efficiency. We, however estimated all models also for the whole sample as a robustness check and the results do not differ in signs or significance levels.} In addition, if the model specified by multisplit lasso includes a binary indicator which represents a category of a nominal variable, we include also all other categories of this variable among the set of predictors. Besides individual \textit{t}-tests we perform an \textit{F}-test of joint significance of the categories of the nominal variable. Similarly as in the case of multisplit lasso, we repeat the resampling and OLS estimation $100$-times. Each time we perform \textit{t}-tests and also a joint \textit{F}-test for each group of dummy variables representing one categorical variable. The \textit{p}-values of the \textit{t}-tests are then aggregated in the same way as in case of multisplit lasso (see above). Further, we calculate mean and median of \textit{p}-values of each joint \textit{F}-test over the $100$ subsamples and according to these statistics we determine whether the dummy indicators of the particular categorical variable should be included. It turns out that every time when a dummy variable representing a category of a nominal variable is chosen by a multiple lasso, both average and median \textit{p}-values of the corresponding \textit{F}-tests are below the significance level (${\alpha=0.05}$). Hence, we include the dummy variables for categories of each nominal variable selected by lasso (see Section~\ref{Results}). % parada. vsechny f testy jsem otestovala u vsech modelu v tabulkach. 27.8. vse v poradku,takze super

@@

One of our goals is to investigate effects of behavioural variables on climate knowledge and concerns about climate change. The behavioural variables that we consider in our study are social value orientation, time preferences, risk preferences, and attitudes towards inequality.

@@

where $P_O$ is the pay-off given to the other party, $P_S$ is the pay-off taken by the player herself and $N$ is the number of games played (in our case $6$).

As one may notice in Table~\ref{PotentialPredictors} in Appendix~$3$, we also include dummy variables for four types of social value orientation (i.e. altruist, prosocial, individualist, competitive) among potential predictors in the lasso estimators. These types are defined based on ring measure~\eqref{RingMeasure}. Each dummy variable corresponds to one of four non-overlapping intervals of the values of ring measure~\eqref{RingMeasure}. None of these dummy variables was selected by lasso into any of our models, therefore we do not discuss them in more detail.

As a basic measure of time preferences we use derived annual discount rates (in percentage), for investing now for one year from now and we refer to this variable as 'Discount rate year from now' in the present study. To obtain the data which would allow us to infer the discount rates, the respondents played games and answered questions informed by \citet{voors2012violent}, \citet{ifcher2011} and \citet{tanaka2010riskTime}. How the time preferences were derived is described in \cite{SurveyUK}. Besides using discount rates for investing now and getting returns in one year, we also implied other types of discount rates.  These are discount rates for $(i)$~ investing now for getting returns in five years $(ii)$~investing in one year for getting returns in two years from now and $(iii)$~ investing in one year from now for getting return in six years from now. None of them was found to be significant, thus we do not further discuss them.

We use two parameters which describe inequity aversion (\citealp{bergson1938, bergson1954, samuelson1956}), in particular the rate of inequity aversion and the subsistence or reserve income. 
To infer these parameters, respondents were choosing from various distributions of income between three hypothetical people. One of them was higher on average but more unequal and the other was lower on average but more equal. The respondents were asked two sets of choice questions. In one of them, the income distribution was centred on the $70^{th}$ percentile of the UK income distribution and in the other the distribution was centred on $40^{th}$ percentile of the UK income distribution. Given the respondents' answers to the two choice-sets, the rate of inequity aversion and subsistence was obtained for each respondent based on equations~\eqref{InequityEq}:

\begin{equation}\label{InequityEq}
\begin{array}{lcl}

\sum_{i=1}^{3}\frac{(Y^H_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^H_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma} \\

\vspace{0.2cm}\\
\sum_{i=1}^{3}\frac{(Y^L_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^L_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma}
\end{array}
\end{equation}

where $\gamma$ is the rate of inequity aversion, $\underline{Y}$ is the subsistence or reserve income and $Y^H_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $70^{th}$ percentile of the UK income distribution. Analogously, $Y^L_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $40^{th}$ percentile of the UK income distribution. To obtain the inequity parameters, equations~\eqref{InequityEq} were solved for $\gamma$ and $\underline{Y}$ while minimizing distance of $\underline{Y}$ to zero.

In theory, the rate of inequity aversion is a continuous measure. However, we consider it as a categorical one as in our dataset it is equal to one of $16$ distinct values for each respondent.\footnote{We also estimated variants of models where the rate of inequity aversion is considered as scale for completeness but we do not present them to save space. However, the results do not differ substantially from those presented here.} These $16$ values and the corresponding frequencies can be found in Table~\ref{FreqiencyCat2} in Appendix~$3$. The subsistence parameter was not selected by lasso into any of our models thus we do not discuss it in more detail.

To test significance of risk aversion, we use various risk aversion coefficients which were estimated for each person from four different utility functions using Bayesian inference \citep{Balcombe2015}. The utility functions are power, logarithmic, exponential and quadratic and we use the estimates of their means and medians. None of them is significant or chosen by lasso in any of our models. For the economy of space we only present models with median or mean of power function. The estimates are very similar when we use other risk aversion coefficients.

As the behavioural variables are not significant in our study, we described their measures only briefly. For more detailed description see \citet{SurveyUK}. The descriptive statistics of these variables (except of inequity aversion) can be found in Table~\ref{Descriptive} in Appendix~$3$. As explained above, we consider the inequity aversion rate as a categorical variable and its frequencies are in Table~\ref{FreqiencyCat2} in Appendix~$3$.

@@

Table~\ref{Know435} summarizes estimates of the predictors of climate change knowledge which we found to be important by means of multisplit lasso estimator. In particular, three predictors are chosen by lasso (see first column in Table~\ref{Know435}). Total score on financial literacy is the number of correct answers out of three finance related mathematical problems \citep{SurveyUK}.\footnote{We also consider answers to each of the three problems separately as individual potential predictors. Two of them are labelled understands inflation and understands compound interest and they are identified as important predictors in other models later.} However, when we re-estimate the model using jackknife OLS with all relevant dummy variables, all categories of total score on financial literacy are insignificant. Furthermore, the model suffers from multicollinearity as the coefficient of correlation between cognitive reflection and total score of financial literacy is equal to $0.343$ and its \textit{p}-value is smaller than $2\times10^{-8}$. For illustration, estimates of jackknife OLS with all explanatory variables listed in Table~\ref{Know435} including financial literacy are shown in Table~\ref{Know435se} in Appendix~$3$. The last column of Table~\ref{Know435se} includes variance inflation factors (VIF) which confirm the presence of multicollinearity. Because of the multicollinearity and insignificance of total score on financial literacy we do not further consider this variable as a predictor of climate knowledge.

The estimates of jackknife OLS without the financial literacy are summarised in the last two columns of Table~\ref{Know435}. The other two variables which were found to be important in explaining climate knowledge are gender and cognitive reflection test~\citep{Frederick2005}. We use the latter as a measure of numeracy and ability of analytical reasoning.

The cognitive reflection test is fully described in~\citet{Frederick2005} and it consists of three numerical problems. The value of our variable is the number of correct answers out of the three questions.\footnote{The possible values are integers and half-integers between zero and three including zero and three as we also recognise if respondent solves half of a problem. Hence, if a respondent answers for example one and half problems correctly, her score is~$1.5$.} The frequencies of values of this variable are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$. To account for plausible non-linear relationship between the test score and cognitive ability we treat the variable as categorical with the base category zero. As it is apparent from Table~\ref{Know435}, the respondents who solved all three problems correctly have significantly higher level of climate knowledge compared to those who did not solve any of them. According to the jackknife OLS, climate knowledge is on average higher also for respondents who answered two problems correctly. However, the effect is larger for three correctly answered problems. Expectedly, the effect of numeracy is positive.

@@

We now discuss the estimates of the model with dependent variable which answers the question whether respondent feels to be more affected by climate policy ($0$) or by climate change ($10$). The estimates are shown in Table~\ref{Climpol435}. We can see that the selected predictors are climate knowledge, understanding of inflation and risk assessment consistency.

The level of understanding of inflation is based on respondents' answer to the following numerical problem: 'Imagine that the interest rate on your savings account was 1 percent per year and
inflation was 2 percent per year. After 1 year, would you be able to buy...' The respondents should choose one of the three following answers: $(i)$~'More than today with the money in
this account' $(ii)$~'Exactly the same as today with the money in
this account' $(iii)$~'Less than today with the money in
this account'. Correct answer is 'less'. The value of the variable is equal to one if the respondent answers 'less', it is equal to $0.5$ if she answers 'the same' and it is equal to zero if she answers 'more'. The frequencies of answers are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$. Since it is quite possible that the effect of our measure of understanding of inflation is non-linear we treat the variable as categorical with the base category zero.

As it is apparent from Table~\ref{Climpol435}, understanding of inflation and risk assessment consistency increase the likelihood of being subjectively more affected by climate policy than by climate change. This is likely to be because the two predictors are highly correlated with financial literacy. The correlation coefficient of understanding of inflation and financial literacy is $0.694$ and the correlation coefficient of risk assessment consistency and financial literacy is $0.145$. Both correlation coefficients are highly significant with \textit{p}-value lower than $2.00\times 10^{-8}$.~~\footnote{Risk assessment consistency is a binary variable so we also run a two sample \textit{t}-test to measure correlation between risk assessment consistency and financial literacy. In particular, we applied a two sample \textit{t}-test to test if mean financial literacy is statistically equal for the respondent who answered risk questions consistently and for those with inconsistent answers to risk questions. The test statistic is highly significant with \textit{p}-value lower than $2.00\times 10^{-8}$. Hence, the mean financial literacy is different in these two groups which is in accordance with the significance of the correlation coefficient} It is intuitive, that the respondents with higher level of financial literacy are more likely to see how their wealth and way of living can be affected by climate policy through environmental tax rates.

It was previously shown that interactions of measures of cognitive ability and ideological and political world-view are strong predictors of attitudes towards climate change rather than cognitive ability or numeracy itself \citep{Kahan2012, KahanEtAl2012, Hamilton2011, HamiltonKeim2009}. In accordance with this (as we discuss in more detail in Section~\ref{ResWTP} below) we detect a significant impact of interactions of an indicator of political and cultural world-view and a measure of numeracy (and ability of analytical, technical reasoning) on WTP for climate change mitigation. Therefore, we also estimate variants of the models presented in this section with the interaction terms included among the predictors but we did not find them to be significant for climate risk perception. We do not present the results in our study to keep its length within reasonable limits.\footnote{It would probably be more revealing to test for significance of interactions of cognitive ability and political orientation, but unfortunately, the respondents were not asked about their political or partisan preferences directly in the survey.}

\vspace{1.5cm}

{\centering
\begin{threeparttable}
\caption{\small{\textit{\textbf{Climate versus policy effects perception:} Multisplit lasso and jackknife OLS}}}
% soubor LassoClimpol43 Climpol435.R
\label{Climpol435} 
\centering
\begin{small}
\begin{tabular}{lclccl} 
\hline \vspace{-0.2cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Climate knowledge&$1.0000$& &$0.2158$&$1.0000$&\\
  \\
\vspace{-0.2cm}Climate knowledge - squared&$<2.00\times 10^{-8}$&$^{***}$&$-0.0607$&$0.0124$&$^{*}$\\
  \\
\vspace{-0.2cm}Understands inflation $=0.5$&$1.0000$& &$-0.0394$&$1.0000$& \\
  \\
\vspace{-0.2cm}Understands inflation $=1$&$0.0130$&$^{*}$&$-0.5759$&$6.27\times 10^{-6}$&$^{***}$\\
    \\
  Consistent answers to risk&\multirow{2}{*}{$1.06\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-0.5885$}&\multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
\hspace{0.6cm}questions $(0/1)$&& &&&\\
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
\begin{tablenotes}
  \begin{footnotesize}
   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
  \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
     \end{adjustwidth}
 \singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

As we can see in Tables~\ref{Climcare435} and~\ref{Climpol435}, the linear climate knowledge term is positive (and insignificant), but the squared term is negative and significant for both climate seriousness and climate change versus policy. That means, the effect of climate knowledge is positive but decreasing for low levels of knowledge, while for medium and high degree of climate knowledge the effect is negative. The negative effect of higher levels of climate knowledge may seem to be counter-intuitive, however it is less surprising in the light of previous literature. According to \citet{reynolds2010now}, people who know less about climate change tend to ascribe unrealistic consequences (for example skin cancer) to global warming. The authors argue that it is even possible for any future ecological or political disaster to be viewed as a consequence of global warming by public \citep{WhatPplKnow94}. Also the level of familiarity with causes and basic mechanisms of climate change is quite unsatisfactory. For example, members of public tend to confuse climate and weather\footnote{Actually, some researchers confuse climate and weather too \citep{Hsiang2013, DesGreenstone2007}.} and as a result they mostly agree with the statement that climate changes from year to year \citep{reynolds2010now}. It is therefore understandable that individuals who are weaker in climate knowledge are more concerned about climate change and its consequences.

Another possible explanation of the negative and significant climate knowledge effect can be affective orientation towards global warming \citep{Kahan2015}. It was previously shown that individuals, who believe that climate change is real and caused by humans and who correctly assert, for instance, that usage of fossil fuels is one of the causes of global warming are also likely to affirm other, perhaps false statements which are consistent with higher environmental risks \citep{reynolds2010now}. An example of such a false statement is that atmospheric emissions of sulphur contribute to global warming \citep{Kahan2015}. The OCSI questions which we use to measure climate knowledge are true/false statements and more than half of them is of a same type as the sulphur emissions statement above. That is, the correct answer does not evince concerns about climate change while the incorrect one does. This could explain why the respondents who believe that climate change is quite serious are likely to score lower on climate knowledge. If this is true, the climate concern variables are predictors of climate knowledge and not the other way around. Therefore, if this is true, climate knowledge should not be included in the specifications with estimates summarised in Tables~\ref{Climcare435} and~\ref{Climpol435}. As a robustness test, we estimate the models for climate seriousness perception and climate change versus climate policy not including the climate knowledge as an explanatory variable. In both cases, the estimates of the rest of the explanatory variables and their significance levels are almost the same as in the case with climate knowledge and they are summarised in Table~\ref{ClimCare435eClimpol435f} in Appendix~$3$.

\subsection{Willingness to pay for climate change mitigation}\label{ResWTP}
This section is focused on models explaining preferred gas and electricity tax rates, which we use as a measure of WTP for climate change mitigation. The estimates of the multisplit lasso and the jackknife OLS are summarised in Table~\ref{Gas435cp}.

One of the important selected predictors is age. The age was recorded as a categorical variable with the lowest category $24$ or younger, the second lowest category is $25-34$, the third one is $35-44$ and so on up to the highest category which is  $75$ or older. We use the lowest age group ($24$ or younger) as the base category. Coefficients of all higher categories are negative and with exemption of $35-44$ and $75$ or older they are all significant. Thus, WTP declines with age, perhaps because older people have lower likelihood of experiencing tougher consequences of climate change predicted for more distant future \citep{Hamilton2011}.\footnote{We also estimated the model with interactions of age and number of children and grandchildren as older people who have more offspring can obviously be more concern about future than those who do not have children. However, we did not find the interactions to be significant. We do not present the results in this study to save space.}

{\centering
\begin{threeparttable}
\caption{\small{\textit{\textbf{WTP climate - gas and electricity tax:} Multisplit lasso and jackknife OLS}}}
% soubor Gas435cp.R, LassoGas47b.R
% soubor Gas435cp2.R, LassoGas47b.R

%jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation
\label{Gas435cp} 
\centering
\begin{small}
\begin{tabular}{lclccl} 
\hline \vspace{-0.35cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}} %LassoGas47b
  & \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\%Gas435cp

\vspace{-0.35cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.35cm}Age\tnote{a} $~25-34$&$1\times10^{-6}$&$^{***}$&$-13.271$&$0.266$& \\
        \\
\vspace{-0.35cm}Age $35-44$&$0.006$&$^{**}$&$-29.494$&$3\times10^{-7}$&$^{***}$\\
        \\ 
 \vspace{-0.35cm}Age $45-54$&$1.000$& &$-34.625$&$<2\times10^{-8}$&$^{***}$\\
        \\ 
  \vspace{-0.35cm}Age $55-64$&$1.000$& &$-40.061$&$<2\times10^{-8}$&$^{***}$\\
        \\       
 \vspace{-0.35cm}Age $65-74$&$1.000$& &$-46.006$&$<2\times10^{-8}$&$^{***}$\\
        \\ 
 \vspace{-0.35cm}Age$~75$ or older&$1.000$& &$-26.071$&$1.000$& \\
        \\       
 Climate versus policy&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$10.408$}&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
\vspace{-0.35cm}\hspace{0.6cm}effects perception&&&\\
 \\

  \vspace{-0.35cm}Inequity aversion (categorical)\tnote{a}&\multicolumn{2}{c}{\textit{negative cor.~}$^{*}$}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\
      \\
 Equal intergenerational&\multirow{2}{*}{$0.011$}&\multirow{2}{*}{${*}$}&\multirow{2}{*}{$20.760$}&\multirow{2}{*}{$0.002$}&\multirow{2}{*}{$^{**}$}\\
\vspace{-0.35cm}\hspace{0.6cm}allocation of resources $(0/1)$\tnote{b}&&&\\
 \\  
 Understands compound&\multirow{2}{*}{$1.000$}&\multirow{2}{*}{ }&\multirow{2}{*}{$-2.942$}&\multirow{2}{*}{$1.000$}&\multirow{2}{*}{ }\\
\vspace{-0.35cm}\hspace{0.6cm}interest $=0.5$&&&\\
 \\   
 Understands compound&\multirow{2}{*}{$1\times10^{-5}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-39.381$}&\multirow{2}{*}{$3\times10^{-5}$}&\multirow{2}{*}{$^{***}$}\\
\vspace{-0.35cm}\hspace{0.6cm}interest $=1$&&&\\
  \\  
\vspace{-0.35cm}Understands inflation $=0.5$&$1.000$& &$-15.892$&$0.516$& \\
  \\
\vspace{-0.35cm}Understands inflation $=1$&$6\times10^{-5}$&$^{***}$&$-42.711$&$<2\times10^{-8}$&$^{***}$\\
  \\
 Consistent answers to risk&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-34.861$}&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
 \hspace{0.6cm}  \vspace{-0.35cm} questions $(0/1)$&& &&&\\
\\ Consistent answers within&\multirow{2}{*}{$0.045$}&\multirow{2}{*}{$^{*}$}&\multicolumn{3}{c}{\textit{Not included}}\\
\hspace{0.6cm}investments $(0/1)$\tnote{c}&& &&&
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}

     \item[~]\textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
\vspace{-0.4cm}  \begin{adjustwidth}{0.65cm}{} \item \vspace{-0.1cm}The signs of the significant coefficients in the multisplit lasso are same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.\vspace{0.05cm}
\item[a]Age is only available as categorical with  base category '$24$ or younger'. Inequity aversion treated as categorical (see Section~\ref{BehaviouralVars}).
\vspace{0.05cm}
\item[b]This variable is equal to $1$ for those respondents who believe that their income and standard of living generally is about equal as the income and standard of living of their parents (when they were about the respondent's age) and it is also equal to the income and standard of living of their children (when they will reach the respondent's age). The variable is equal to $0$ for all other respondents.
\vspace{0.05cm}
\item[c]We eventually excluded this variable from the further analysis. Although consistency within investment was selected by lasso,  it is only marginally significant and strongly correlated with risk assessment consistency. Furthermore, even after exclusion of this variable the model includes relatively large number of predictors and their signs and significance levels do not change.
    \end{adjustwidth}

\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

We can see in Table~\ref{Gas435cp} that another very strong predictor of WTP is perception of climate versus policy effects on ones way of living which we analyse as a dependent variable in Section~\ref{ResPerc}. Expectedly, those who perceive effects of climate change as more serious than effects of climate policy have higher WPT for climate change mitigation. As we will discus below, this variable is a partial mediator of impact of financial literacy.

Another predictor of WTP is inequity aversion. As we explained in Section~\ref{BehaviouralVars}, we treat this variable as categorical. The values of our inequity aversion measure and their frequencies are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$. Although the effect of inequity aversion is largely positive and decreasing, the signs and significance levels vary across the categories without any clear pattern.

It is further apparent from Table~\ref{Gas435cp} that WTP is higher for respondents who believe that their income and standard of living generally is about equal to the income and standard of living of their parents when they were about the respondent's age and they also believe that their income and standard of living is equal to the income and standard of living of their children when they will reach the respondent's age.

>>>>

\large Findings:
\begin{itemize}
\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important

\item Mar-Sept last year temperature very important for maize yields
\item SD temperature last year positive and significant
\item dry spell 20 MAM last year important (but not dry spell MAM10)
\item interesting. Precipitation 2 months MAM last year very significant and positive
\item mean temp last year negative and significant, hill shaped 
\end{itemize}

\underline{New findings:}

\begin{itemize}
\item The yields seem to be more responsive to weather on west than on east
\end{itemize}
\normalsize
\FloatBarrier
\section{Methodology}\label{Method}

@@

@@

@@

@@

@@
